---
title: "Project trial"
author: "Dipankar Lahiri"
date: "2025-03-29"
output: html_document
---

This is a trial run of an upcoming data analysis project. Made-up data is being used.
**Research Question**: Do people who use AI more come from higher parental income families, have lower academic scores, worse social skills, and use digital social media platforms more?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(moments)

df <- read_csv('mock_data.csv')


```

# Research Design
**Data Collection**: Survey, Questionnaire.  
**Population**: Young Chinese people (18-25)  
**Sample Size**: Approximately 50 (convenience or self-selecting)

# Variables
- **Dependent variable**: AI use (frequency of use)  
- **Independent variables**:
  1. Academic performance (self-reported grades)  
  2. Parental income (categorical income ranges)  
  3. Social interaction skills (composite score from multiple Likert scale questions)  
  4. Social media exposure (frequency or duration of social media use)

# Hypotheses
- **H1**: Lower academic performance is associated with higher AI use.  
- **H2**: Higher parental income is associated with higher AI use.  
- **H3**: Worse social interaction skills are associated with higher AI use.  
- **H4**: Higher social media exposure is associated with higher AI use.

# Ordering Data, Summary Statistics

```{r}

df <- df %>%
  mutate(
    `AI Use (Frequency)` = factor(`AI Use (Frequency)`, levels = c("Never", "Rarely", "Sometimes", "Often", "Always")),
    `Academic Performance (Grade)` = factor(`Academic Performance (Grade)`, levels = c("F", "D", "C", "B", "A")),
    `Parental Income (Range)` = factor(`Parental Income (Range)`, 
                                       levels = c("< ¥3,000", "¥3,000 - ¥5,000", "¥5,001 - ¥8,000", "¥8,001 - ¥12,000", "> ¥12,000"),
                                       ordered = TRUE),
    `Field of Study` = factor(`Field of Study`)
  )

cat("\n=== Summary of Dataset ===\n")
summary(df)

cat("\n=== AI Use (Frequency) ===\n")
print(table(df$`AI Use (Frequency)`))
print(prop.table(table(df$`AI Use (Frequency)`)))

cat("\n=== AI Use (Purpose) ===\n")
print(prop.table(table(df$`AI Use (Purpose)`)))
cat("Most Common Purpose: ", names(sort(table(df$`AI Use (Purpose)`), decreasing = TRUE))[1], "\n")

cat("\n=== Academic Performance (Grade) ===\n")
print(prop.table(table(df$`Academic Performance (Grade)`)))

cat("\n=== Parental Income (Range) ===\n")
print(prop.table(table(df$`Parental Income (Range)`)))

cat("\n=== Social Interaction Skills (Score) ===\n")
print(c(
  Mean = mean(df$`Social Interaction Skills (Score)`),
  SD = sd(df$`Social Interaction Skills (Score)`),
  Median = median(df$`Social Interaction Skills (Score)`),
  IQR = IQR(df$`Social Interaction Skills (Score)`),
  Skewness = skewness(df$`Social Interaction Skills (Score)`),
  Kurtosis = kurtosis(df$`Social Interaction Skills (Score)`)
))

cat("\n=== Social Media Exposure (Hours/day) ===\n")
print(c(
  Mean = mean(df$`Social Media Exposure (Hours/day)`),
  SD = sd(df$`Social Media Exposure (Hours/day)`),
  Median = median(df$`Social Media Exposure (Hours/day)`),
  IQR = IQR(df$`Social Media Exposure (Hours/day)`),
  Skewness = skewness(df$`Social Media Exposure (Hours/day)`),
  Kurtosis = kurtosis(df$`Social Media Exposure (Hours/day)`)
))

cat("\n=== Most Common Field of Study ===\n")
cat(names(sort(table(df$`Field of Study`), decreasing = TRUE))[1], "\n")

```

# Hypothesis Testing

## H1: Lower academic performance is associated with higher AI use
(Null Hypothesis: There is no association between academic performance and AI use.)

```{r}


```

Spearman or Kendall

## H2: Higher parental income is associated with higher AI use.
(Null Hypothesis: There is no association between parental income and AI use.)

```{r}


```

Spearman, Kendall

## H3: Worse social interaction skills are associated with higher AI use.
(Null Hypothesis: There is no association between social interaction skills and AI use.)

```{r}

boxplot(df$`Social Interaction Skills (Score)` ~ df$`AI Use (Frequency)`, 
        xlab = "AI Use Frequency", ylab = "Social Interaction Skills (Score)", 
        main = "Boxplot of Social Interaction Skills by AI Use Frequency")

df$`AI Use (Frequency) Ranked` <- as.numeric(rank(df$`AI Use (Frequency)`))
df$`Social Interaction Skills (Score) Ranked` <- as.numeric(rank(df$`Social Interaction Skills (Score)`))

cor.test(df$`Social Interaction Skills (Score) Ranked`, df$`AI Use (Frequency) Ranked`, method = "spearman")

cor.test(df$`Social Interaction Skills (Score) Ranked`, df$`AI Use (Frequency) Ranked`, method = "kendall")


```




## H4: Higher social media exposure is associated with higher AI use.
(Null Hypothesis: There is no association between social media exposure and AI use.)




```{r}


```



